The Practical Application of Computerised Multistage Tests

### 2021-02-26 article.Rmd
Added scaling function

### Converting True Score to Theta

@@

### Module A Mapping Function, $g_{A1}()$

First we use a polynomial regresssion to fit a set of coefficients to the Rasch model.

@@

# use a quintic polynomial regression model to make predictions
lm_moduleA_scores_poly <- lm(theta ~ poly(trueScore, degree=5, raw=TRUE), data = df_moduleA_TS)

# setup the score scale for module A
moduleA_TSRange <- data.frame(
  trueScore = c(seq(0,20)),
  theta = c(rep(0, 21))
)

# plot the theta/true score curve from the Rasch model
plot(x=df_moduleA_TS$theta, y=df_moduleA_TS$trueScore, 
     type="l", col = "coral1", main="Module A", xlab="θ", ylab="True Score")

# overlay the polynomial mapped true score points
scaleScores_moduleA_poly <- predict(lm_moduleA_scores_poly, moduleA_TSRange)
moduleA_TSRange$theta <- unname(scaleScores_moduleA_poly)
points(x=moduleA_TSRange$theta, y=moduleA_TSRange$trueScore, col="cornflowerblue", pch=4)

legend(2, 9, legend=c("MCC", "Quintic Poly Function"),
  col=c("coral1", "cornflowerblue"), lty=1:2, cex=0.8)

summary(lm_moduleA_scores_poly)
```

@@

```{r Module A True Score Function}
options(scipen = 1, digits = 3)
gA_coeffs <- unname(coef(lm_moduleA_scores_poly))

@@

The Avellain developers, when they want to calculate the ability for a candidate could use such a function to map from the number correct (NC) score to an estimated theta for candidate $i$: $\hat\theta_i$. For example, using R:

@@

### Validating the Mapping Function
We can now use this function to overlay the original MCC:

@@

# plot the theta/true score curve from the Rasch model
plot(x=df_moduleA_TS$theta, y=df_moduleA_TS$trueScore, 
     type="l", col = "coral1", main="Module A", xlab="θ", ylab="True Score")
moduleA_TSRange$theta <- gA1(moduleA_TSRange$trueScore)
points(x=moduleA_TSRange$theta, y=moduleA_TSRange$trueScore, col="purple", pch=4)
legend(2, 9, legend=c("MCC", "gA1()"),
  col=c("coral1", "purple"), lty=1:2, cex=0.8)
```
We can see that it is a good fit, both in the middle of the range, and importantly at the upper and lower bounds.

### Module D-H Mapping Functions, $g_{D1}(), g_{E1}(), g_{F1}, g_{G1}(), g_{H1}()$

```{r}
df_moduleD_TS <- trueScoreSummary(plot(raschmodel, type = "ICC", items = moduleD1_items, plot = FALSE))
lm_moduleD_scores_poly <- lm(theta ~ poly(trueScore, degree=5, raw=TRUE), data = df_moduleD_TS)
gD_coeffs <- unname(coef(lm_moduleD_scores_poly))

df_moduleE_TS <- trueScoreSummary(plot(raschmodel, type = "ICC", items = moduleE1_items, plot = FALSE))
lm_moduleE_scores_poly <- lm(theta ~ poly(trueScore, degree=5, raw=TRUE), data = df_moduleE_TS)
gE_coeffs <- unname(coef(lm_moduleE_scores_poly))

df_moduleF_TS <- trueScoreSummary(plot(raschmodel, type = "ICC", items = moduleF1_items, plot = FALSE))
lm_moduleF_scores_poly <- lm(theta ~ poly(trueScore, degree=5, raw=TRUE), data = df_moduleF_TS)
gF_coeffs <- unname(coef(lm_moduleF_scores_poly))

df_moduleG_TS <- trueScoreSummary(plot(raschmodel, type = "ICC", items = moduleG1_items, plot = FALSE))
lm_moduleG_scores_poly <- lm(theta ~ poly(trueScore, degree=5, raw=TRUE), data = df_moduleG_TS)
gG_coeffs <- unname(coef(lm_moduleG_scores_poly))

df_moduleH_TS <- trueScoreSummary(plot(raschmodel, type = "ICC", items = moduleH1_items, plot = FALSE))
lm_moduleH_scores_poly <- lm(theta ~ poly(trueScore, degree=5, raw=TRUE), data = df_moduleH_TS)
gH_coeffs <- unname(coef(lm_moduleH_scores_poly))
```

```{r Output g Functions}
panel1_coeffs <- data.frame(t(gA_coeffs))
panel1_coeffs <- rbind(panel1_coeffs, data.frame(t(gD_coeffs)))
panel1_coeffs <- rbind(panel1_coeffs, data.frame(t(gE_coeffs)))
panel1_coeffs <- rbind(panel1_coeffs, data.frame(t(gF_coeffs)))
panel1_coeffs <- rbind(panel1_coeffs, data.frame(t(gG_coeffs)))
panel1_coeffs <- rbind(panel1_coeffs, data.frame(t(gH_coeffs)))
panel1_coeffs$moduleID <- c('A1', 'D1', 'E1', 'F1', 'G1', 'H1')
panel1_coeffs <- panel1_coeffs[, c(7, 6, 5, 4, 3, 2, 1)]

# OUTPUT a CSV file too...
write_csv(panel1_coeffs, "../data/scoring-functions-panel1.csv")
```

@@

### Item Information Functions (IIFs)

@@

### Module Information Functions (MIFs)

@@

## Routing and Cut-Scores

Routing between modules can then determined by cut scores derived from the intersection of information functions for the modules in the next stage. We should aim to place the cut score to be that at which the MIFs cross.

@@

We can then repeat the exercise for the stage 3 modules.

```{r Stage 3 MIFs}
plotData <- rbind(
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleD1_items, plot = FALSE), 
                          "D"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleE1_items, plot = FALSE), 
                          "E"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleF1_items, plot = FALSE), 
                          "F"))

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Information Function", x="θ", y="I(θ)")

plotData <- rbind(
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleE1_items, plot = FALSE), 
                          "E"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleF1_items, plot = FALSE), 
                          "F"),
  informationCoefficients(plot(raschmodel, 
                               type = "IIC", 
                               items = moduleG1_items, plot = FALSE), 
                          "G"))

ggplot(plotData, aes(x=theta, y=information)) + 
  geom_line(aes(colour=module)) +
  labs(title="Module Information Function", x="θ", y="I(θ)")
```

>>>>

#### Converting True Score to Theta

@@

#### Module A Mapping Function, $g_{A}()$

We'll use a polynomial regression to fit a set of coefficients to the MCC.

@@

# use a quintic polynomial regression to model the MCC
lm_moduleA_scores_poly <- lm(theta ~ poly(trueScore, degree=5, raw=TRUE), data = df_moduleA_TS)

# show a summary of the polynomial regression
summary(lm_moduleA_scores_poly)
```

@@

```{r Module A True Score Function}
gA_coeffs <- unname(coef(lm_moduleA_scores_poly))

@@

The test development team, when they want to calculate the ability for a candidate could use such a function to map from the number correct (NC) score to an estimated theta for candidate $i$: $\hat\theta_i$. For example, using an R function this would be:

@@

#### Validating the Mapping Function

We can now use this function to overlay the original MCC:

@@

# plot the theta/true score curve from the Rasch model
plot(x=df_moduleA_TS$theta, y=df_moduleA_TS$trueScore, 
     type="l", col = "coral1", main="Module A", xlab="θ", ylab="True Score")
moduleA_TSRange$theta <- gA1(moduleA_TSRange$trueScore)
points(x=moduleA_TSRange$theta, y=moduleA_TSRange$trueScore, col="purple", pch=4)
legend(2, 9, legend=c("MCC", "gA()"),
  col=c("coral1", "purple"), lty=1:2, cex=0.8)
```
We can see that it is a good fit, both in the middle of the range, and importantly at the upper and lower bounds.

We then need to go through each module in every panel and create a true score mapping function for each. These functions, or the lookup-tables that they can be used to generate, will then need to be provided to the developers of the test delivery software so that they can be used to calculate theta estimates and scale scores for each module.

#### Scale Scoring Function, $s()$

A scale score function can then be created to map the $\hat{\theta}$ to a score that will be reported back to the candidate and the testing institute. Typically a summary of capability such as a CEFR rating will also be provided with the scale score.

The test administrators may want to create different scale score functions for each stage of the MST, although the simplest approach is to just have a single one. A straightforward scaling function to use is to simply normalise the ability estimate for a candidate to a range 0-1 and then apply a multiplier constant $k$:

$$
s(\hat{\theta}_i) = k\frac{\hat{\theta}_i - min(\theta)}{max(\theta) - min(\theta)}
$$

Usually the scale score would be rounded to a whole number. For example, in R the scaling function might look like this:

```{r}
s <- function(thetaEstimate, thetaRange) {
  k <- 100
  round((thetaEstimate - thetaRange[1]) / (thetaRange[2] - thetaRange[1]) * k)
}

scaleScore <- s(thetaEstimate = -0.5, thetaRange = c(-4, 4))

(scaleScore)
```

More complex scaling functions that use additional terms can also be used to tailor the reporting score range to the requirements of the test institute.

### Cut scores for routing

@@

#### Item Information Functions (IIFs)

@@

#### Module Information Functions (MIFs)

@@

Routing between modules can then determined by cut scores derived from the intersection of information functions for the modules in the next stage. We should aim to place the cut score to be that at which the MIFs intersect.

@@
